---
title: "Aula 8"
output: html_notebook
---

# AULA 8 - Econometria II

Objetivos:

1.  Dar continuidade à Introdução ao Software Estatístico R e pacotes econométricos disponíveis.
2.  Estimação e Interpretação dos Modelos com dados em Painel Multiperíodos: Estimação em Primeiras Diferenças

Referências:

1.  Wooldridge J.M. (2012) "Introductory Econometrics", p.412.
2.  Kleiber, C. & Zeileis, A. (2008) "Applied Econometrics with R", p. 84.
3.  Croissant, Y. & Millo, G. (2019) "Panel Data Econometrics with R by", p.1.

Sites que contém material auxiliar sobre a aula:

-   <http://eclr.humanities.manchester.ac.uk/index.php/Panel_in_R>
-   <https://justinmshea.github.io/wooldridge/articles/Introductory-Econometrics-Examples.html#chapter-13-pooling-cross-sections-across-time-simple-panel-data-methods>
-   <https://justinmshea.github.io/wooldridge/index.html>

## Bibliotecas

```{r}
source("R/libraries.R")
```

## County crime rates in North Carolina

### Dados

Data on 90 counties in North Carolina, for the years 1981 through 1987. Various factors including geographical location, attitudes toward crime, historical records, and reporting conventions might be contained in a~i~.

-   **crmrte** is number of crimes per person.
-   **prbarr** is the estimated probability of arrest
-   **prbconv** is the estimated probability of conviction (given an arrest)
-   **prbpris** is the probability of serving time in prison (given a conviction)
-   **avgsen** is the average sentence length served
-   **polpc** is the number of police officers per capita.

We use the logs of all variables to estimate elasticities. We also include a full set of year dummies to control for state trends in crime rates. We can use the years 1982 through 1987 to estimate the differenced equation

```{r}
crime4 <- tibble::as_tibble(wooldridge::crime4)

head(crime4)
```

### Modelo

```{r}
model_1 <- plm::plm(
  lcrmrte ~ d83 + d84 + d85 + d86 + d87 + lprbarr + lprbconv + lprbpris + lavgsen + lpolpc, 
  data = crime4, 
  effect = "individual", 
  model = "fd", # first differences 
  index = c("county", "year")) # identificadores de observações (cross section e temporal)

summary(model_1)
```

Obs: A função plm para opção `fd` calcula o modelo com dummies em primeiras diferenças; não altera as estimaões dos efeitos parciais.
